---
title: "Assassination Classroom"
author: "Ben Kelley"
date: "2024-11-05"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading
### All necessary packages and dependencies are at the top of this chunk.

```{r}
library(jsonlite)
library(sf)
library(sp)
library(dlm)
library(hms)

# data can be found in Professor Lee's Coursworks for STAT 3105.

dat <- read_json('~/gps/20200819132607.geojson')
head(dat[['features']], 1)
```

## Data Processing

```{r}
# I will extract the Latitude, Longitude, Date, and Time from the geojson file and reorganize it into an easily readable data frame

timestamps <- c()

for (i in 1:length(dat$features)){
  
  timestamps[i] <- dat$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(dat$features)){
  
  longitude[i] <- dat$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(dat$features)){
  
  latitude[i] <- dat$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords <- coordinates(utm_df)

# Remove the last row of utm_coords to fix a dimension issue with the filter later on
utm_coords <- utm_coords[-249, ]

head(utm_coords)
```

## Filter Construction

### What is the average time between records?

```{r}
head(data$Timestamp, 20)

# It appears to be about 30 seconds, so dt will be 30
```

```{r}
gps_variance <- 20^2
v_mat <- matrix(c(abs(rnorm(1, 0, gps_variance)), 0, 0, 0,
                  0, abs(rnorm(1, 0, gps_variance)), 0, 0,
                  0, 0, abs(rnorm(1, 0, gps_variance)), 0,
                  0, 0, 0, abs(rnorm(1, 0, gps_variance))), byrow = T, ncol = 4)
dt <- 30
g_mat <- matrix(c(1, 0, dt, 0,
                  0, 1, 0, dt,
                  0, 0, 1, 0,
                  0, 0, 0, 1), byrow=TRUE, ncol=4)
avg_walk_speed_m_per_sec <- 1.4  # https://en.wikipedia.org/wiki/Walking
dlm_spec <- dlm(
  FF= matrix(c(1, 0, 0, 0,
               0, 1, 0, 0,
               0, 0, 1, 0,
               0, 0, 0, 1), byrow = T, ncol = 4),
  GG= g_mat,
  V = v_mat,
  W = diag(c(5, 5, 1, 1)^2),
  m0 = matrix(c(utm_coords[1, ], rep(avg_walk_speed_m_per_sec / dt, 2)),
              ncol=1), # A vector by R defaults is a k by 1 matrix
  C0 = diag(rep(10^2, 4)))

d1dlm_filter_mod <- dlmFilter(utm_coords, dlm_spec)
```

## Plot the Data

```{r}
plot(d1dlm_filter_mod$y, xlab = "Longitude", ylab = "Latitude")
```
## Upload and Plot the Other 10 Files
```{r}
d2 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200818114606.geojson')

timestamps <- c()

for (i in 1:length(d2$features)){
  
  timestamps[i] <- d2$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d2$features)){
  
  longitude[i] <- d2$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d2$features)){
  
  latitude[i] <- d2$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords2 <- coordinates(utm_df)

d2dlm_filter_mod <- dlmFilter(utm_coords2, dlm_spec)
```


```{r}
d3 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200820151044.geojson')

timestamps <- c()

for (i in 1:length(d3$features)){
  
  timestamps[i] <- d3$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d3$features)){
  
  longitude[i] <- d3$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d3$features)){
  
  latitude[i] <- d3$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords3 <- coordinates(utm_df)

d3dlm_filter_mod <- dlmFilter(utm_coords3, dlm_spec)
```

```{r}
d4 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200821111447.geojson')

timestamps <- c()

for (i in 1:length(d4$features)){
  
  timestamps[i] <- d4$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d4$features)){
  
  longitude[i] <- d4$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d4$features)){
  
  latitude[i] <- d4$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords4 <- coordinates(utm_df)
utm_coords4 <- utm_coords4[-667, ]

d4dlm_filter_mod <- dlmFilter(utm_coords4, dlm_spec)
```

```{r}
d5 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200824130857.geojson')

timestamps <- c()

for (i in 1:length(d5$features)){
  
  timestamps[i] <- d5$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d5$features)){
  
  longitude[i] <- d5$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d5$features)){
  
  latitude[i] <- d5$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords5 <- coordinates(utm_df)

d5dlm_filter_mod <- dlmFilter(utm_coords5, dlm_spec)
```

```{r}
d6 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200825121346.geojson')

timestamps <- c()

for (i in 1:length(d6$features)){
  
  timestamps[i] <- d6$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d6$features)){
  
  longitude[i] <- d6$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d6$features)){
  
  latitude[i] <- d6$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords6 <- coordinates(utm_df)
utm_coords6 <- utm_coords6[-711, ]

d6dlm_filter_mod <- dlmFilter(utm_coords6, dlm_spec)
```

```{r}
d7 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200826131614.geojson')

timestamps <- c()

for (i in 1:length(d7$features)){
  
  timestamps[i] <- d7$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d7$features)){
  
  longitude[i] <- d7$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d7$features)){
  
  latitude[i] <- d7$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords7 <- coordinates(utm_df)
utm_coords7 <- utm_coords7[-865, ]

d7dlm_filter_mod <- dlmFilter(utm_coords7, dlm_spec)
```

```{r}
d8 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200827113234.geojson')

timestamps <- c()

for (i in 1:length(d8$features)){
  
  timestamps[i] <- d8$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d8$features)){
  
  longitude[i] <- d8$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d8$features)){
  
  latitude[i] <- d8$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords8 <- coordinates(utm_df)

d8dlm_filter_mod <- dlmFilter(utm_coords8, dlm_spec)
```

```{r}
d9 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200828122627.geojson')

timestamps <- c()

for (i in 1:length(d9$features)){
  
  timestamps[i] <- d9$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d9$features)){
  
  longitude[i] <- d9$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d9$features)){
  
  latitude[i] <- d9$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords9 <- coordinates(utm_df)
utm_coords9 <- utm_coords9[-267, ]

d9dlm_filter_mod <- dlmFilter(utm_coords9, dlm_spec)
```

```{r}
d10 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200828130816.geojson')

timestamps <- c()

for (i in 1:length(d10$features)){
  
  timestamps[i] <- d10$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d10$features)){
  
  longitude[i] <- d10$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d10$features)){
  
  latitude[i] <- d10$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords10 <- coordinates(utm_df)
utm_coords10 <- utm_coords10[-391, ]

d10dlm_filter_mod <- dlmFilter(utm_coords10, dlm_spec)
```

```{r}
d11 <- read_json('C:/Users/btkre/OneDrive/Documents/Stat 3105/gps/20200831115147.geojson')

timestamps <- c()

for (i in 1:length(d11$features)){
  
  timestamps[i] <- d11$features[[i]]$properties$time
}

longitude <- c()

latitude <- c()

for (i in 1:length(d11$features)){
  
  longitude[i] <- d11$features[[i]]$geometry$coordinates[[1]]
}

for (i in 1:length(d11$features)){
  
  latitude[i] <- d11$features[[i]]$geometry$coordinates[[2]]
}


# Let's convert our latitude and longitude to meters

data <- data.frame(Longitude = longitude, Latitude = latitude, Timestamp = timestamps)
longlat <- data.frame(Longtiude = longitude, Latitude = latitude)
timer <- data.frame(Time = timestamps)
# This following step is similar to transforming time stamps like "2020-01-25" into a time object
spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   # This needs to be a data frame
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))
# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords11 <- coordinates(utm_df)
utm_coords11 <- utm_coords11[-867, ]

d11dlm_filter_mod <- dlmFilter(utm_coords11, dlm_spec)
```

## Plot all of the days together to observe patterns

```{r}
plot(d1dlm_filter_mod$y, xlab = "Longitude", ylab = "Latitude", col = 1)
points(d2dlm_filter_mod$y, col = 2)
points(d3dlm_filter_mod$y, col = 3)
points(d4dlm_filter_mod$y, col = 4)
points(d5dlm_filter_mod$y, col = 5)
points(d6dlm_filter_mod$y, col = 6)
points(d7dlm_filter_mod$y, col = 7)
points(d8dlm_filter_mod$y, col = 8)
points(d9dlm_filter_mod$y, col = 9)
points(d10dlm_filter_mod$y, col = 10)
points(d11dlm_filter_mod$y, col = 11)
```

# Date Predicting Function

*For this function, I've examined the start times of each day of the week and decided that the best way to predict which day of the week the input data will be from is to see which day, from the first week, has the most similar start time. This function makes that prediction.*

```{r}

time_date <- function(second, minute, hour){
  
  predicted_day <- c()
  
  if (hms(second, minute, hour) <= hms(00, 34, 17)){
    
    predicted_day <- "Friday"
  }
  
  else if (hms(second, minute, hour) > hms(00, 34, 17) & hms(second, minute, hour) <= hms(30, 30, 18)){
    
    predicted_day <- "Tuesday"
  }
  
  else if (hms(second, minute, hour) > hms(30, 30, 18) & hms(second, minute, hour) <= hms(34, 19, 19)){
    
    predicted_day <- "Monday"
  }
  
  else if (hms(second, minute, hour) > hms(34, 19, 19) & hms(second, minute, hour) <= hms(00, 20, 20)){
    
    predicted_day <- "Wednesday"
  }
  
    else {
    
    predicted_day <- "Thursday"
    }
  
  return(predicted_day)
}
```
## Filter Choosing Function

```{r}
filter_choice <- function (character_vec) {

  if (character_vec == "Monday") {
  return(d5dlm_filter_mod)
}
else if (character_vec == "Tuesday"){
  
  return(d2dlm_filter_mod)
}
  else if (character_vec == "Wednesday") {
    
    return(d1dlm_filter_mod)
  }
  
  else if (character_vec == "Thursday") {
    
    return(d3dlm_filter_mod)
  }
  
  else {
    
    return(d4dlm_filter_mod)
  }
}
```

## Step Choosing Function

*The algorithm will attempt to plant the bomb 7 minutes into Professor Lee's commute. Since each file collects data at a different rate, it is necessary to choose a different step count depending on which file the algorithm chooses.*

```{r}
step_choice <- function (character_vec) {

  if (character_vec == "Monday") {
    
  return(57)
}
else if (character_vec == "Tuesday"){
  
  return(14)
}
  else if (character_vec == "Wednesday") {
    
    return(18)
  }
  
  else if (character_vec == "Thursday") {
    
    return(27)
  }
  
  else {
    
    return(34)
  }
}
```

## The Bomb Plan

```{r}
bomb_location <- function(start_long, start_lat, hours, minutes, seconds) {
  
  start_time <- hms( seconds, minutes, hours)
  
day_of_week <- time_date(second = seconds, minute = minutes, hour = hours)

pred_filter <- filter_choice(day_of_week)

longlat <- data.frame(Longtiude = start_long, Latitude = start_lat)

timer <- data.frame(Time = start_time)

spat_df <- SpatialPointsDataFrame(coords= longlat,
                                  data=timer,   
                                  proj4string=CRS("+proj=lonlat +datum=WGS84"))

# This step converts the longitude/latitude -> UTM
utm_df <- spTransform(spat_df, CRSobj = "+proj=utm +zone=12 +datum=WGS84")
utm_coords <- coordinates(utm_df)

coords <- (pred_filter$y)

index <- 1:nrow(coords)

long_mod <- lm(coords[, 1] ~ index)

pred_steps <- step_choice(day_of_week)

long_pred_data <- data.frame(coords = utm_coords[1], index = pred_steps)

predicted_longitude <- predict(long_mod, long_pred_data)

lat_mod <- lm(coords[, 2] ~ index)

lat_pred_data <- data.frame(coords = utm_coords[2], index = pred_steps)

predicted_latitude <- predict(lat_mod, lat_pred_data)

new_min <- minutes + 7

predicted_time <- hms(seconds, new_min, hours)

cat("Bomb Longitude (Meters):", predicted_longitude, "     " ,"Bomb Latitude (Meters):", predicted_latitude, "     " ,"Time:")
print(predicted_time)
}
```

## Test

*I have chosen coordinates and a time similar to one of the days in the first week just to test the algorithm. Any starting data may be entered instead.*

```{r}
bomb_location(start_long = -114, start_lat = 47, seconds = 55, minutes = 30, hours = 19)
```

## Additional Information
### How does temperature affect the speed of walking?

The temperature data was taken from the following source: https://world-weather.info/forecast/usa/new_york/august-2020/

```{r}
tempmat <- matrix(c(mean(d1dlm_filter_mod$f),
mean(d2dlm_filter_mod$f),
mean(d3dlm_filter_mod$f),
mean(d4dlm_filter_mod$f),
mean(d5dlm_filter_mod$f),
mean(d6dlm_filter_mod$f),
mean(d7dlm_filter_mod$f),
mean(d8dlm_filter_mod$f),
mean(d9dlm_filter_mod$f),
mean(d11dlm_filter_mod$f),
82,
73,
79,
82,
88,
91,
79,
90,
86,
73), ncol = 2, byrow = F)

tempmod <- lm(tempmat[,1] ~ tempmat[, 2])
summary(tempmod)
```

This model gives a positive coefficient to the effect of temperature, meaning that the target walks faster when the temperature is warmer.

## Model Validation

```{r}
coords <- d3dlm_filter_mod$y
index <- 1:nrow(d3dlm_filter_mod$y)

val_mod <- lm(coords[, 1] ~ index)

plot(index, coords[, 1], main = "Day 3 Longitude Model")
abline(a = val_mod$coefficients[1], b = val_mod$coefficients[2], col = "red")

plot(val_mod$residuals)
abline(h = 0, col = "red")

coords <- d5dlm_filter_mod$y
index <- 1:nrow(d5dlm_filter_mod$y)

val_mod2 <- lm(coords[, 2] ~ index)

plot(index, coords[, 2], main = "Day 5 Latitude Model")
abline(a = val_mod2$coefficients[1], b = val_mod2$coefficients[2], col = "red")

plot(val_mod2$residuals)
abline(h = 0, col = "red")

coords <- d2dlm_filter_mod$y
index <- 1:nrow(d2dlm_filter_mod$y)

val_mod3 <- lm(coords[, 1] ~ index)

plot(index, coords[, 1], main = "Day 1 Longitude Model")
abline(a = val_mod3$coefficients[1], b = val_mod3$coefficients[2], col = "red")

plot(val_mod3$residuals)
abline(h = 0, col = "red")
```

We can see here that the data is not linear or Normal, and sometimes is heteroscedastic as well. This dramatically undermines confidence in the models' predictive powers.
